Disclosure of Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy

### 2018-04-23 aggregation/agg-agg.tex
Reviewer response
Late nite work
Revision 3

\subsubsection{Unit Process Inventory}

A unit process makes up a single shared column of $A_f$, $A_d$, and $B_f$, as depicted in Figure~\ref{fig:aggregation}a.  Documentation of unit processes is specified in ISO 14048, and several compliant formats exist, the most well-known being the ILCD and Ecospold XML formats.  ISO 14048 specifies a distinction between intermediate and elementary flows, but does not include the concept of a model foreground.  In a matrix representation, each unit process must have exactly one reference flow, which corresponds to its row and column in $A_f$.

@@

See Figure~\ref{fig:aggregation}b.  A foreground can be viewed as a linked set of unit processes.  The collection of processes themselves make up the foreground of the study, and so exchanges among them are recorded in $A_f$, and exchanges outside the foreground are represented in $A_d$ or $B_f$.  A complete foreground represents
the full disclosure of a model, and a reader can replicate and extend it independently.
%An excerpt from a larger study may also be published as a complete foreground. %
Any modeling technique used to treat a multi-functional process, such as an allocation or system expansion, can only be fully expressed as a foreground with at least as many columns as outputs.  For examples of multi-output processes represented as foreground models, please see the supporting information.  The example shown in Figure~\ref{fig:aggregation}b shows five nodes, including one cut-off flow

@@

See Figure~\ref{fig:aggregation}d.  Using this approach, the
 aggregated dependency vector is split into two parts that sum to the original:
\begin{equation}
 \tilde{\mathbf{a}}_d = \tilde{\mathbf{a}}_{d,priv} + \tilde{\mathbf{a}}_d'
\end{equation}
The disclosed dependencies $\tilde{\mathbf{a}}_d'$ are reported, and the private dependencies $\tilde{\mathbf{a}}_{d,priv}$ are replaced with an aggregated background inventory derived from the background database, $\tilde{\mathbf{b}}_{x,agg}$.  This approach can be used when a reader lacks access to the background inventory sets referred to in $\tilde{\mathbf{a}}_{d,priv}$, or when the author wishes not to disclose the private dependencies.  Foreground emissions $\tilde{\mathbf{b}}_f$ can be reported separately from the aggregation result, but often in current practice they are not distinguished from the background flows.

>>>>

\subsubsection{Single-Output Unit Process}

A single-output unit process makes up a single shared column of $A_f$, $A_d$, and $B_f$, as depicted in Figure~\ref{fig:aggregation}a.  Documentation of unit processes is specified in ISO 14048, and several compliant formats exist, the most well-known being the ILCD and Ecospold XML formats.  ISO 14048 specifies a distinction between intermediate and elementary flows, but does not include the concept of a model foreground.  In a matrix representation, each unit process must have exactly one reference flow, which corresponds to its row and column in $A_f$.
Please see the supporting information for more detail on preparing the $A_f$ matrix.

@@

See Figure~\ref{fig:aggregation}b.  A foreground can be viewed as a linked set of unit processes.  The collection of processes themselves make up the foreground of the study, and so exchanges among them are recorded in $A_f$, and exchanges outside the foreground are represented in $A_d$ or $B_f$.  A complete foreground represents
the full disclosure of a model, and a reader can replicate and extend it independently.
%An excerpt from a larger study may also be published as a complete foreground. %
Any modeling technique used to treat a multi-functional process, such as an allocation or system expansion, can only be fully expressed as a foreground with at least as many columns as outputs.  For examples of multi-output processes represented as foreground models, please see the supporting information.  The example shown in Figure~\ref{fig:aggregation}b shows five nodes, including one cut-off flow.

@@

See Figure~\ref{fig:aggregation}d.  Using this approach, the
 aggregated dependency vector is split into two parts that sum to the original:
\begin{equation}
 \tilde{\mathbf{a}}_d = \tilde{\mathbf{a}}_{d,priv} + \tilde{\mathbf{a}}_d'
\label{eqn:partial}
\end{equation}
The disclosed dependencies $\tilde{\mathbf{a}}_d'$ are reported, and the private dependencies $\tilde{\mathbf{a}}_{d,priv}$ are replaced with an aggregated background inventory derived from the background database, $\tilde{\mathbf{b}}_{x,agg}$.  This approach can be used when a reader lacks access to the background inventory sets referred to in $\tilde{\mathbf{a}}_{d,priv}$, or when the author wishes not to disclose the private dependencies.  Foreground emissions $\tilde{\mathbf{b}}_f$ can be reported separately from the aggregation result, but often in current practice they are not distinguished from the background flows.

### 2018-04-23 aggregation/agg-basic-lca.tex
Revision 3

In the most general terms, the LCA problem concerns the interaction of a collection of activities and commodities described by a supply-and-use inventory, in which $U$ is a table of the use of each commodity by each activity (inputs) and $V$ is a table of the supply of each commodity by each activity (outputs) \citep{Pauliuk_2015_framework}.  Such an account can always be transformed into a \textit{symmetric} table of inputs and outputs on a commodity-by-commodity or industry-by-industry basis \citep{Eurostat_2008}. Doing so requires the application of various assumptions and transformations to the data, the details of which are not in scope of this paper, but which have recently been reviewed in depth \citep{Suh_JIE_2010, Majeau_Bettez_2014}.

A symmetric input output table can be written as a table of coefficients, in which each column of the matrix provides a ``recipe'' for creating the corresponding commodity by consuming other commodities (and, if applicable, producing other co-products), equivalent to a ``unit process'' in ISO LCA, except that the reference flow is implicit in the column's index.  The table of coefficients, commonly known as a \textit{direct requirements matrix}, can be used to compute the output from all activities that is necessary to produce a given commodity.  That vector of outputs can then be combined with a second table of coefficients, known as an \textit{emission matrix}, that reports exchanges with the environment associated with the production of each commodity, to determine the cumulative environmental emissions associated with the production of the commodity over its life cycle.  Those emissions are individually \textit{characterized} with respect to an environmental impact category of interest, and the characterized emissions summed to report the \textit{impact score} or category indicator, which is the result of the study.  The formulation may be summarized:
\begin{equation}
s = \mathbf{c}^T \cdot B \cdot \left(I - A\right)^{-1} \cdot \mathbf{y}
\label{eqn:leontief}
\end{equation}
Here $A$ represents the direct requirements; $B$ is the emission matrix;  $\mathbf{c}$ is a column vector of characterization factors for the environmental emissions; $\mathbf{y}$ is the externally specified final demand; and $s$ is the numerical impact score or category indicator.  For simplicity, we will consider the computation of a single impact result.  However, it is straightforward to imagine the more typical case, in which $C$ is a matrix of characterization vectors, and $\mathbf{s}$ is a vector of results.  Eq.~\ref{eqn:leontief} is visualized in Figure~\ref{fig:lca}.
\input{fig_lca}

The term $(I-A)$ is equivalent to a normalized \textit{technology matrix}, and in certain cases can be computed from the difference between the supply and use tables, $(V'-U)$, with equivalent results \citep{Suh_JIE_2010}. 
The coefficient matrices $A$ (or $I-A$) and $B$ together are called a \textit{life cycle inventory database} (LCIDB), and the functional unit of a study is given by the final demand vector $\mathbf{y}$, which is the product or service whose delivery is to be assessed.\footnote{%
Differences in convention have led to both the direct requirements matrix $A$, and the technology matrix, analogous to $(I-A)$, being denoted by the same symbol in different contexts.  If the data contained in the LCIDB is sufficiently rich, either approach may be implemented with equivalent computational results.  The direct requirement formulation has the advantage of permitting iterative solution, which can dramatically reduce the computational resources required \citep{Peters_JLCA_2007}.}  Intermediate results of this computation include:
\begin{itemize}
\item $\mathbf{x} = \left(I-A\right)^{-1}\cdot\mathbf{y}$, the activity level vector;
\item $\mathbf{b} = B\cdot\mathbf{x}$, the life cycle inventory;
\item $s = \mathbf{c}^T\cdot\mathbf{b}$, the life cycle impact category indicator, the study result.
\end{itemize}

Some LCA software systems have the capacity to export the $A$ and $B$ matrices for a given study model.  However, the contents of commercial LCIDBs are proprietary and subject to licensing restrictions, and moreover are not usually modified by study authors.  Thus exporting $A$ and $B$ does not meet our authority and primacy objectives.  Additionally, replicating the entire LCIDB is cumbersome because of the large size of the matrices.   As a consequence, the matrix formulation in Eq.~\ref{eqn:leontief} is unsuitable as a disclosure framework.

>>>>

In the most general terms, the LCA problem concerns the interaction of a collection of activities and commodities described by a supply-and-use inventory, in which $U$ is a table of the use of each commodity by each activity (inputs) and $V$ is a table of the supply of each commodity by each activity (outputs) \citep{Pauliuk_2015_framework}.  Such an account can always be transformed into a \textit{symmetric} table of inputs and outputs on a commodity-by-commodity or industry-by-industry basis \citep{Eurostat_2008}. Doing so requires the application of various assumptions and transformations to the data, the details of which are not in scope of this paper, but which have recently been reviewed in depth \citep{Majeau_Bettez_2014}.  Although they originated independently, the ``classical'' process-based approach and the input-output approach have been shown to be equivalent for the purposes of LCA computation \citep{Suh_JIE_2010}.  Please see the Supporting information for a discussion about this equivalency.

A symmetric input output table can be written as a table of coefficients, in which each column of the matrix provides a ``recipe'' for creating the corresponding commodity by consuming other commodities (and, if applicable, producing other co-products), equivalent to a ``unit process'' in ISO LCA, except that the process must have exactly one reference flow, the identity of which is implicit in the column's index.  The table of coefficients, commonly known as a \textit{direct requirements matrix}, can be used to compute the output from all activities that is necessary to produce a given commodity.  That vector of outputs can then be combined with a second table of coefficients, known as an \textit{emission matrix}, that reports exchanges with the environment associated with the production of each commodity, to determine the cumulative environmental emissions associated with the production of the commodity over its life cycle.  Those emissions are individually \textit{characterized} with respect to an environmental impact category of interest, and the characterized emissions summed to report the \textit{impact score} or category indicator, which is the result of the study.  The formulation may be summarized:
\begin{equation}
s = \mathbf{c}^T \cdot B \cdot \left(I - A\right)^{-1} \cdot \mathbf{y}
\label{eqn:leontief}
\end{equation}
Here $A$ represents the direct requirements, sometimes known as the \textit{Leontief matrix}; $B$ is the emission matrix;  $\mathbf{c}$ is a column vector of characterization factors for the environmental emissions; $\mathbf{y}$ is the externally specified final demand; and $s$ is the numerical impact score or category indicator.  For simplicity, we will consider the computation of a single impact result.  However, it is straightforward to imagine the more typical case, in which $C$ is a matrix of characterization vectors, and $\mathbf{s}$ is a vector of results.  Eq.~\ref{eqn:leontief} is visualized in Figure~\ref{fig:lca}.
\input{fig_lca}

The term $(I-A)$ is equivalent to a \textit{technology matrix} in traditional LCA, normalized by column so that each column represents a unit output of a single reference product, and each row represents the use of the corresponding product. 
The coefficient matrices $A$ (or $I-A$) and $B$ together are called a \textit{life cycle inventory database} (LCIDB), and the functional unit of a study is given by the final demand vector $\mathbf{y}$, which is the product or service whose delivery is to be assessed.
%%\footnote{Differences in convention have led to both the direct requirements matrix $A$, and the technology matrix, analogous to $(I-A)$, being denoted by the same symbol in different contexts.  If the data contained in the LCIDB is sufficiently rich, either approach may be implemented with equivalent computational results.  The direct requirement formulation has the advantage of permitting iterative solution, which can dramatically reduce the computational resources required \citep{Peters_JLCA_2007}.}
Intermediate results of this computation include:
\begin{itemize}
\item $\mathbf{x} = \left(I-A\right)^{-1}\cdot\mathbf{y}$, the activity level vector;
\item $\mathbf{b} = B\cdot\mathbf{x}$, the life cycle inventory;
\item $s = \mathbf{c}^T\cdot\mathbf{b}$, the life cycle impact category indicator, the study result.
\end{itemize}

Some LCA software systems have the capacity to export the $A$ and $B$ matrices for a given study model.  However, the contents of commercial LCIDBs are proprietary and subject to licensing restrictions, and moreover are not usually modified by study authors.  Thus exporting $A$ and $B$ does not meet our authority and originality objectives.  Additionally, replicating the entire LCIDB is cumbersome because of the large size of the matrices.   As a consequence, the matrix formulation in Eq.~\ref{eqn:leontief} is unsuitable as a disclosure framework.

### 2018-04-23 aggregation/agg-disclosure.tex
Revision 3

A disclosure containing these six elements satisfies most of the requirements established at the top of this section (the exception is ensuring the privacy of confidential data, which is treated later in the article):
\begin{itemize}
  \item The result is \textbf{computable} using Eq.~\ref{eqn:study};
  \item the computation is \textbf{complete} and \textbf{minimal} because all components are required and no superfluous components are included;
  \item The results are \textbf{reproducible} if the reader is able to accurately interpret the author's references in \ref{itm:bg} and \ref{itm:em};
  \item The disclosure includes only information for which the author can claim \textbf{authority};
  \item The disclosure includes all information specific to the study, or for which the study is the \textbf{primary} source, including re-implementations of already-published data. 
\end{itemize}
As mentioned, the proper interpretation of the disclosure requires that the reader accurately interprets the author's references to background databases and elementary flows, including mutually consistent implementation of the LCIA vector $\mathbf{c}$.  These aspects are out of scope for the purposes of this paper, and we will assume that a shared understanding between author and reader can be reached.  However, the challenges of operationalizing this requirement are complex (see the discussion and supporting information).  %Some implications on the information required for a successful disclosure are considered in the Discussion.

>>>>

A disclosure containing these six elements satisfies most of the requirements established at the top of this section (the exception is ensuring the privacy of confidential data, which is treated later in the article):
\begin{itemize}
  \item The result is \textbf{computable} using Eq.~\ref{eqn:study};
  \item the computation is \textbf{complete} and \textbf{minimal} because all components are required and no superfluous components are included;
  \item The results are \textbf{reproducible} if the reader is able to accurately interpret the author's references in \ref{itm:bg} and \ref{itm:em};
  \item The disclosure includes only information for which the author can claim \textbf{authority};
  \item The disclosure includes all information specific to the study, or for which the study is the \textbf{original} source, including re-implementations of already-published data. 
\end{itemize}
As mentioned, the proper interpretation of the disclosure requires that the reader accurately interprets the author's references to background databases and elementary flows, including mutually consistent implementation of the LCIA vector $\mathbf{c}$.  These aspects are out of scope for the purposes of this paper, and we will assume that a shared understanding between author and reader can be reached.  However, the challenges of operationalizing this requirement are complex (see the discussion and supporting information).  %Some implications on the information required for a successful disclosure are considered in the Discussion.

### 2018-04-23 aggregation/agg-lca-study.tex
Revision 3

This same workflow can be repeated to build a more elaborate foreground.  For instance, a model of private vehicle travel may make use of reference processes for vehicle production and decommissioning, road maintenance, fuel retail, and a model of direct emissions from fuel combustion.  This model could occupy an additional column of $\tilde{A}$, with the coefficients for fuel combustion emissions being contained in the corresponding column of $\tilde{B}$.  A process for brewing coffee could be created in a third column that requires coffee beans, water, and electricity. Finally, a fourth column could be added that invokes the first three columns in proportion to describe the activity of traveling to a coffeeshop to work.%\footnote{No prizes for guessing where this passage was written.}.

@@

Without loss of generality, it is possible to construct the study such that the \textit{first} column of $\tilde{A}$ delivers the functional unit of the study.  We thereby designate the vector $\tilde{\mathbf{y}} =  [ 1,\, 0,\, 0 ,\,\ldots,\, 0]^{T}$ as the \textit{canonical functional unit} of a foreground study, which allows us to introduce convenient notation later on.

>>>>

This same workflow can be repeated to build a more elaborate foreground.  For instance, a model of private vehicle travel may make use of reference processes for vehicle production and decommissioning, road maintenance, fuel retail, and a model of direct emissions from fuel combustion.  This model could occupy an additional column of $\tilde{A}$, with the coefficients for fuel combustion emissions being contained in the corresponding column of $\tilde{B}$.  A process for brewing coffee could be created in a third column that requires coffee beans, water, and electricity. Finally, a fourth column could be added that invokes the products from the first three columns in proportion to describe the reference product of traveling to a coffeeshop to work.%\footnote{No prizes for guessing where this passage was written.}.

@@

Without loss of generality, it is possible to construct the study such that the \textit{first} column of $\tilde{A}$ delivers the functional unit of the study.  We thereby designate the vector $\tilde{\mathbf{y}} =  [ 1,\, 0,\, 0 ,\,\ldots,\, 0]^{T}$ as the \textit{canonical functional unit} of a foreground study, which allows us to introduce convenient notation later on.\footnote{Of course, the ``functional unit'' does not need to be limited to a unit of any given output---instead, it \textit{defines} a unit of output. For instance, if the functional unit of a study is 1,000 loaves of bread, and one loaf of bread weighs 0.454 kg, and the unit output from column 567 of the A matrix is 1 kg of bread, then the canonical functional unit would be realized by entering \texttt{454} in column 1, row 567, leaving the rest of that column zero.}

### 2018-04-23 aggregation/agg-multioutput.tex
Revision 3

According to MWS14, Partitioning allocation (PA) is a ``production-balanced model that artificially splits the requirements of an activity across its different outputs.''  The exact manner of the partitioning need not be specified, as long as the parts sum up to the whole.  In other words, for a process with $n$ outputs the modeler has $n-1$ degrees of freedom in specifying the allocated inventories, with the requirement that the $n$th inventory equals the remainder.

In the disclosure framework, any manner of production-balanced partitioning can be represented by including the raw, un-allocated inventory as one foreground column, choosing any output as the reference.  The other $n-1$ outputs are then included as arbitrarily-specified allocated inventories in additional columns, for a total of $n$ columns.  Each column must be normalized to a unit reference.  The reference flow not explicitly allocated will simply be computed as the remainder when the allocated inventories are subtracted from the un-allocated inventory.

@@

Using the foreground in Table~\ref{tbl:partition}, LCA computation using the canonical functional unit will automatically compute the self-consistent single-output result for 1 kg of chlorine gas because the allocated inventories for hydrogen and sodium hydroxide will be subtracted from the total.  The foreground activity levels for the system are:

@@

In the classical system expansion case, applying the expanded functional unit of 1 kg chlorine, plus 0.028 kg hydrogen and 1.13 kg of sodium hydroxide, will result in a unit activity of the reference process, without any cut-off flows..  Equivalently, the canonical functional unit, production of 1 kg of chlorine gas, will result in the same process activity level, and will also generate the co-products as cutoff flows.

>>>>

According to MWS14, Partitioning allocation (PA) is a ``production-balanced model that artificially splits the requirements of an activity across its different outputs.''  The exact manner of the partitioning need not be specified, as long as the parts sum up to the whole.  In the disclosure framework, any manner of production-balanced partitioning can be represented by including the raw, un-allocated inventory as one foreground column, choosing any output as the reference.  The other $n-1$ outputs are then included as allocated inventories in additional columns, for a total of $n$ columns.  Each column must be normalized to a unit reference.

Note that, although it is customary for allocated process inventories to be derived as a uniform partition of the un-allocated process based on some parameter such as mass, price or energy content, this is not necessary.  For a process with $n$ outputs the modeler has $n-1$ degrees of freedom in specifying the allocated inventories, with the requirement that the $n$th inventory equals the remainder.  In the framework presented, the modeler must provide the $n-1$ allocated inventories explicitly.  The reference flow for the $n$th process will simply be computed as the remainder when the allocated inventories are subtracted from the un-allocated inventory.

@@

Using the foreground in Table~\ref{tbl:partition}, LCA computation using the canonical functional unit (see main text, Eq.~\ref{eqn:foreground}) will automatically compute the self-consistent single-output result for 1 kg of chlorine gas because the allocated inventories for hydrogen and sodium hydroxide will be subtracted from the total.  The foreground activity levels for the system are:

@@

In the classical system expansion case, applying the expanded functional unit of 1 kg chlorine, plus 0.028 kg hydrogen and 1.13 kg of sodium hydroxide, will result in a unit activity of the reference process, without any cut-off flows.  Equivalently, the canonical functional unit, production of 1 kg of chlorine gas, will result in the same process activity level, and will also generate the co-products as cutoff flows.

### 2018-04-23 aggregation/agg-requirements.tex
Revision 3

\subsection{Authority and Primacy}

@@

Because of these conditions, in the vast majority of cases the LCA study disclosure must include the author's reimplementations in order to achieve both the aims for authority and primacy.  In so doing, the disclosure enables a critical reviewer to evaluate whether the author's implementations are generally correct and complete.

>>>>

\subsection{Authority and Originality}

@@

Because of these conditions, in the vast majority of cases the LCA study disclosure must include the author's reimplementations in order to achieve both the aims for authority and originality.  In so doing, the disclosure enables a critical reviewer to evaluate whether the author's implementations are generally correct and complete.

### 2018-04-23 aggregation/proofs.tex
Revision 3

Given a life cycle inventory database (LCIDB) containing input-output matrix $A$ and emission matrix $B$, and a final demand vector $\mathbf{y}$, construct an augmented inventory database
$\tilde{A} = \left[\begin{smallmatrix} 0 & \mathbf{0} \\ \mathbf{y} & A \end{smallmatrix}\right]$,
$\tilde{B} = [ \mathbf{0},\, B ]$, and canonical functional unit $\tilde{\mathbf{y}} =  [ 1,\, 0,\, 0 ,\,\ldots,\, 0]^{T}$.  Show that
\begin{equation} B\times\left(I-A\right)^{-1}\times\mathbf{y} = \tilde{B}\times\left(I-\tilde{A}\right)^{-1}\times\tilde{\mathbf{y}}
\end{equation}

We begin by constructing $(I-\tilde{A})^{-1}$.  Because the matrix is block triangular, we only need to determine the value of $\mathbf{q}$ in the matrix below:
\begin{eqnarray}
  I &=&  (I-\tilde{A})^{-1}\times(I-\tilde{A})\\
  \left[\begin{array}{cc} 1 & \mathbf{0} \\
    \mathbf{0} & I
  \end{array}\right] &=&
  \left[\begin{array}{cc} 1 & \mathbf{0} \\
      \mathbf{q} & (I-A)^{-1}
  \end{array}\right]\times
  \left[\begin{array}{cc} 1 & \mathbf{0} \\
      -\mathbf{y} & (I-A)
        \end{array}\right]
\end{eqnarray}
It can be seen that $\mathbf{q}-(I-A)^{-1}\times\mathbf{y}=0$ is a necessary condition to satisfy the equality, and thus $\mathbf{q}=(I-A)^{-1}\times\mathbf{y}$.

>>>>

Given a life cycle inventory database (LCIDB) containing input-output matrix $A$ and emission matrix $B$, and a final demand vector $\mathbf{y}$, construct an augmented inventory database
$\tilde{A} = \left[\begin{smallmatrix} 0 & \mathbf{0}^T \\ \mathbf{y} & A \end{smallmatrix}\right]$,
$\tilde{B} = [ \mathbf{0},\, B ]$, and canonical functional unit $\tilde{\mathbf{y}} =  [ 1,\, 0,\, 0 ,\,\ldots,\, 0]^{T}$.  Show that
\begin{equation} B\times\left(I-A\right)^{-1}\times\mathbf{y} = \tilde{B}\times\left(I-\tilde{A}\right)^{-1}\times\tilde{\mathbf{y}}
\end{equation}

We begin by constructing $(I-\tilde{A})^{-1}$.  Because the matrix is block triangular, we only need to determine the value of $\mathbf{q}$ in the matrix below:
\begin{eqnarray}
  I &=&  (I-\tilde{A})^{-1}\times(I-\tilde{A})\\
  \left[\begin{array}{cc} 1 & \mathbf{0}^T \\
    \mathbf{0} & I
  \end{array}\right] &=&
  \left[\begin{array}{cc} 1 & \mathbf{0}^T \\
      \mathbf{q} & (I-A)^{-1}
  \end{array}\right]\times
  \left[\begin{array}{cc} 1 & \mathbf{0}^T \\
      -\mathbf{y} & (I-A)
        \end{array}\right]
\end{eqnarray}
It can be seen that $\mathbf{q}-(I-A)^{-1}\times\mathbf{y}=\mathbf{0}$ is a necessary condition to satisfy the equality, and thus $\mathbf{q}=(I-A)^{-1}\times\mathbf{y}$.
